A Variant of G\"{o}del's Ontological Proof \\ in a Natural Deduction Calculus \\ (Draft)

### 2014-08-06 Papers/NaturalDeduction/GodProof-ND.tex
Improve introduction of ND paper

G\"odel's notion of positive property and Leibniz's notion of perfection are not exactly the same \citep{Adams}. On a technical level, the main distinction seems to be that G\"odel's positive properties are not just atomic properties, like Leibniz's perfections, but can also consist of complex combinations of atomic properties \citep{fitting}[p.139.] (TODO: Check this claim). In particular, one of G\"odel's axioms states that the conjunction of any set of positive properties is positive. And from this axiom, it follows immediately that the property of being God-like is positive. While this step is intuitively and informally clear, it is not easily formalizable in a typical logical calculus, because it requires inferring that being god-like is a (possibly infinite) conjuntion of positive properties, while it has only been defined as the property holding for individuals who have all positive properties. This interplay between universal quantification (in the definition of a God-like being) and conjunction (in G\"odel's mentioned axiom) is a technical difficulty that probably explains why, starting with Scott \cite{scott}, to whom G\"odel confided his manuscript, this axiom of G\"odel's has been replaced by another one that simply assumes the positivity of the property of being God-like. 
%This axiom is unfortunately equivalent to the possibility of God's existence given the other axioms, leading to criticisms that it might be too strong \citep{sobel2}\citep{fitting}[Section 11.4. Objections.].

@@

This paper presents new natural deduction proofs of (Scott's version of) the lemmas and theorems in G\"odel's ontological argument \cite{Goedel}. In contrast to Scott's version \cite{Scott}, also shown here in natural deduction form, the version presented here is characterized by shorter and more direct derivations.

Section \ref{sec:calculus} defines the higher-order modal natural deduction calculus used to construct the formal proofs. Section \ref{sec:scott} shows Scott's proofs, translated to natural deduction. And section \ref{sec:proof} presents the new proofs.

@@

For Natural deduction

@@

\begin{calculus}
{propositional rules}
{fig:ND}

@@

@@

% \vspace{2em}

% \s\s\s\s
% \infer[\pos_I]{\pos A}{n: \fbox{\infer*{A}{}} }
% \s\s\s\s\s
% \infer[\pos_E]{\eta: \fbox{ \infer*{}{A} } }{\pos A}

@@

\noindent
A \emph{derivation} is a directed acyclic graph whose nodes are formulas and whose edges correspond to applications of the inference rules shown in Figures \ref{fig:ND}, \ref{fig:Classical} and \ref{fig:NDK}. Parts of a derivation may be surrounded by boxes. A \emph{proof} is a derivation that additionally satisfies the following conditions:

\begin{itemize}
\item \textbf{eigen-variable conditions:}
if $\rho$ is a $\all_I$ inference eliminating a variable $\alpha$, then any occurrence of $\alpha$ in the proof should be an ancestor of the occurrence of $\alpha$ eliminated by $\rho$;
if $\rho$ is a $\ex_E$ inference introducing a variable $\beta$, then any occurrence of $\beta$ in the proof should be a descendant of the occurrence of $\beta$ introduced by $\rho$.
%
\item \textbf{eigen-box conditions:} $\omega$ must be a fresh name for a box.
%
\item \textbf{boxed assumption condition:} any assumption should be discharged within the box where it is made.
%
\item \textbf{unboxed root condition:} the proof's root should not be inside any box.
\end{itemize}

>>>>

G\"odel's notion of positive property and Leibniz's notion of perfection are not exactly the same \citep{Adams}. On a technical level, the main distinction seems to be that G\"odel's positive properties are not just atomic properties, like Leibniz's perfections, but can also consist of complex combinations of atomic properties \citep{fitting}[p.139.] (TODO: Check this claim). In particular, one of G\"odel's axioms states that the conjunction of any set of positive properties is positive. And from this axiom, it follows immediately that the property of being God-like is positive. While this step is intuitively and informally clear, it is not easily formalizable in a typical logical calculus, because it requires inferring that being god-like is a (possibly infinite) conjuntion of positive properties, while it has only been defined as the property holding for individuals who have all positive properties. This interplay between universal quantification (in the definition of a God-like being) and conjunction (in G\"odel's mentioned axiom) is a technical difficulty that probably explains why, starting with Scott \cite{scott}, to whom G\"odel confided his manuscript, this axiom of G\"odel's has been replaced by another one that simply assumes the positivity of the property of being God-like.

The main criticism against G\"odel's ontological argument is the \emph{modal collapse}, an undesirable consequence of the stipulated axioms. Most recent works \cite{todo} on the ontological argument have focused on proposing modifications of the argument that would not entail a modal collapse. This is discussed in more detail in Section \ref{sec:collapse}.

%This axiom is unfortunately equivalent to the possibility of God's existence given the other axioms, leading to criticisms that it might be too strong \citep{sobel2}\citep{fitting}[Section 11.4. Objections.].

@@

The first contribution of this paper is the detailed formalization of Scott's version \cite{Scott} of G\"odel's ontological argument \cite{Goedel} (as shown in Section \ref{sec:scott}) in a natural deduction calculus (as defined in Section \ref{sec:calculus}). A natural deduction style \citep{gentzen,prawitz} was chosen mainly for three reasons. Firstly, presentations of G\"oedel's proof are typically either informal or formalized in other styles of calculi (e.g. Fitting's tableaux \cite{fitting} or Sobel's sentential modal calculus \cite{sobel2}). Therefore, a formalization in natural deduction is a valuable complement to the existing presentations. Secondly, natural deduction is a familiar style for many people, who can now benefit from a natural deduction presentation of G\"odel's ontological argument. Finally, as natural deduction is the style used by proof assistants such as Coq \cite{coq} and Isabelle \cite{isabelle}, the natural deduction formalizations have been easily checked step-by-step in Coq \cite{coqformalizations}.

The main contribution of the paper, however, is the presentation of new proofs (also in natural deduction style) of the lemmas, theorems and corollaries in G\"odel's manuscript. In contrast to Scott's proofs \cite{Scott}, the proofs presented here are simpler and shorter, as discussed in Section \ref{sec:conclusion}.

@@

The natural deduction calculus used here has standard rules for propositional connectives and quantifiers, as shown in Figures \ref{fig:PropositionalRules} and \ref{fig:QuantifierRules}. 
To deal with classicality \ref{fig:Classical} and \ref{fig:NDK}. These rules

A \emph{derivation} is a directed acyclic graph whose nodes are formulas and whose edges correspond to applications of the inference rules shown in Figures . Parts of a derivation may be surrounded by boxes. A \emph{proof} is a derivation that additionally satisfies the following conditions:

@@

ToDo: check if disjunction rules are used in Goedel's proof. If not, we could define disjunction as we already do for negation and equivalence.

\begin{calculus}
{propositional rules}
{fig:PropositionalRules}

@@

\textbf{eigen-variable conditions:}
if $\rho$ is a $\all_I$ inference eliminating a variable $\alpha$, then any occurrence of $\alpha$ in the proof should be an ancestor of the occurrence of $\alpha$ eliminated by $\rho$;
if $\rho$ is a $\ex_E$ inference introducing a variable $\beta$, then any occurrence of $\beta$ in the proof should be a descendant of the occurrence of $\beta$ introduced by $\rho$.

@@

\s\s\s\s
\infer[\pos_I]{\pos A}{w: \fbox{\infer*{A}{}} }
\s\s\s\s\s
\infer[\pos_E]{\omega: \fbox{ \infer*{}{A} } }{\pos A}

@@

\begin{itemize}
\item 
%
\item \textbf{eigen-box conditions:} $\omega$ must be a fresh name for a box.
%
\item \textbf{boxed assumption condition:} any assumption should be discharged within the box where it is made.
%
\item \textbf{unboxed root condition:} the proof's root should not be inside any box.
\end{itemize}
